Testable normative theories

### 2020-04-15 simple.tex
Started rework with two stages

\begin{document}
\title{Theories of deliberated preference}
\author{Olivier Cailloux}
\affil{Université Paris-Dauphine, PSL Research University, CNRS, LAMSADE, 75016 PARIS, FRANCE\\
	\href{mailto:olivier.cailloux@dauphine.fr}{olivier.cailloux@dauphine.fr}
}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

@@

\section{Motivation}
In decision theory and economy in general, preferences are generally considered as “intuitive”, as originating from an “immediate sensation”, following the terms used by \citet{von_neumann_theory_2004} (vNM) \citep{fishburn_retrospective_1989, cailloux_reasons_2017}. Though appropriate when studying the everyday behavior of individuals, this conception is insufficient to study the positions that individuals adopt towards difficult decision problems after careful thinking. In this article I define the concept of deliberated preferences to fill this gap: the deliberated preferences of an individual about a given decision problem represent the position of that individual towards that problem that is stable facing counter-arguments.

@@

@@

\subsection{Deliberated preference}
The following important definition is required in order to define the deliberated preference of $i$. An argument $\ar$ is decisive for $\phi$, in short, $\phi\text{-dec}$ iff $\forall \ar' \in \allargs: (\ar', \ar) \ileadstost \phi$. 
\footnote{Equivalently: $\ar \ \phidec ⇔ [\ileadstostinv(\phi)](\ar) = \allargs$, and $\exists \ar \ \phidec ⇔ [\ileadstostinv(\phi)]^{-1}(\allargs) ≠ \emptyset$.}
I also say that $\ar$ is $\propsuredec$, $\proppossdec$, or $¬\proppossdec$, when $\phi=(\prop, \text{sure})$, $\phi=(\prop, \text{possible})$, or $\phi=(¬\prop, \text{possible})$, respectively.
\footnote{$\propsuredec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstosst \prop$, $\proppossdec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstopst \prop$.}

@@

@@

\bibliography{simple}
\end{document}

>>>>

\begin{document}
%\title{Theories of deliberated preference}
\title{Testable normative theories}
\author{Olivier Cailloux}
\affil{Université Paris-Dauphine, PSL Research University, CNRS, LAMSADE, 75016 PARIS, FRANCE\\
	\href{mailto:olivier.cailloux@dauphine.fr}{olivier.cailloux@dauphine.fr}
}
\makeatletter
	\hypersetup{
		pdfsubject={Epistemology},
		pdfkeywords={Decision aiding, Decision making, Argumentation}
	}
\makeatother
\maketitle

@@

\section{Motivation}
TODO The model should be: (m, s) leadsto (t, sure) where m can be for example: after having presented s0, s1, s2, present s3, and tell the individual to wait for one week.

In decision theory and economy in general, preferences are generally considered as “intuitive”, as originating from an “immediate sensation”, following the terms used by \citet{von_neumann_theory_2004} (vNM) \citep{fishburn_retrospective_1989, cailloux_reasons_2017}. Though appropriate when studying the everyday behavior of individuals, this conception is insufficient to study the positions that individuals adopt towards difficult decision problems after careful thinking. In this article I define the concept of deliberated preferences to fill this gap: the deliberated preferences of an individual about a given decision problem represent the position of that individual towards that problem that is stable facing counter-arguments.

@@

\section{General normative theories}
A general normative theory is a tuple $(\alltopic, \allargs, I, (\ileadsto, i \in I), \gargs, \gleadsto)$.

TODO: theorist proposes theory $\gamma$. Contains set $\allargs$. The theory must exhibit a limited number of candidates $s_c$ when $s'$ sometimes trumps $s$ and $s'$ sometimes does not trump $s$ (valid for all $i$): that’s falsifiable, and then permits validation of $\eta$. Also, $\gamma$ must exhibit a replacer $s$ of $s_3, s_1$ against $s_2$, or a set of candidate replacers. The replacer (or candidates) may depend on $s_3$ and $s_1$, but not on $i$. This is falsifiable again. And $\gamma$ must fix bounds of width and breadth. And $\gamma$ must indicate how an $s \notin \args_\gamma$ is unnecessary, by exhibiting resistant [not trumped by any decisive argument in $\args_\gamma$] candidates from $\args_\gamma$ such that at least one trump $s$ or replace $s$.

A counter-theory may propose $\args^\text{ext}$ with some arguments not in $\allargs$ (or, more generally, may propose to enlarge the conditions for considering that $s', s \ileadsto t$: if DM has eaten a pancake or has eaten no pancake, must be stable). Then they either can exhibit a plausible argument in $\args^\text{ext}$ not in $\allargs$ (and indeed $\allargs$ is rejected), or not, and then both theories are equivalent, better keep the simpler one.

Thus, $\gamma$ takes positions only on those things true for all $i$.

Then a theorist can propose a mapping $i$ to $\eta_i$. If $\gamma$ is valid, then it is easy to validate $\eta_i$.

More generally, I need a set of possible situations, and an operation defined by the theory $\gamma$, that brings a situation to another situation, that is normatively better for taking a decision (or innocuous?). The operation is: adding an argument. This operation corresponds to forming the pair $(s', s)$: from a situation $s'$, we add the argument $s$ and bring $i$ in the situation $(s', s)$ where she has heard two arguments.

Then the normative criterion is: $t$ in DJ of $i$ when $\exists s \suchthat \forall s': (s', s) \ileadstost^i t$ (the theory provides a part of this guarantee under falsification, claiming for all $i$, the model makes the rest under validation for this $i$).

@@

\subsection{Deliberated preference}
The following important definition is required in order to define the deliberated preference of $i$. An argument $\ar$ is decisive for $\phi$, in short, $\phi\text{-dec}$ iff $\forall \ar' \in \allargs: (\ar', \ar) \ileadstost \phi$. 
\footnote{Equivalently: $\ar \ \phidec ⇔ [\ileadstostinv(\phi)](\ar) = \allargs$, and $\exists \ar \ \phidec ⇔ [\ileadstostinv(\phi)]^{-1}(\allargs) ≠ \emptyset$.}
I also say that $\ar$ is $\propsuredec$, $\proppossdec$, or $¬\proppossdec$, when $\phi=(\prop, \text{sure})$, $\phi=(\prop, \text{possible})$, or $\phi=(¬\prop, \text{possible})$, respectively.
\footnote{$\propsuredec ⇔ \forall \ar': (\ar', \ar) \ileadstosst \prop ⇔ \forall \ar': ¬[(\ar', \ar) \ileadstope ¬\prop]$, $\proppossdec$ ⇔ $\forall \ar': (\ar', \ar) \ileadstopst \prop$. Note that “both” ≠ “we donno”: in case “both”, we know how to prove that both are reasonable options, in the sense that neither is certain. We have excluded some possibilities (and some other model could disagree with the “both” conclusion).}

@@

\commentOC{TODO}
We want to leave “undecided” as an option: the model takes no position, does not know about $(t, ¬t)$. But we want to exclude $t$ possible and $¬t$ unknown. Thus $t$ possible should imply $¬t$ possible or $t$ sure. If this holds, we can simplify the status of the propositions: only determine whether $t$ is possible for all $t \in \alltopic$, and from there, we know everything. This result can be obtained with: assume $s$ decisive for $t$ poss and no $s$ is decisive for $¬t$ poss, then, for some $(s_1, s)$ leading to $t$ sure and $(s_2, s)$ leading to $\set{t, ¬t}$, it is possible to form $(s_2, s_1, s)$, and if this leads to $t$ sure, it should imply that $s + s_1$ is decisive for $t$ sure; and if this leads to $¬t$ possible, then this should imply that $s+s_2$ is decisive for $¬t$ possible.

@@
